---
title: "ADNI Mixed Effect Models"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(HLSGUtils)
```

# Mixed Effect Models

`wgs_adno_partitioner` function create data partitions from ADNI datasets.
To run mixed effect model on ADNI data we have three data source:

- ADNI dataset
- ADNI WGS dataset
- PCA of ADNI WGS data set

We also set the variables from each data source in function arguments:
- clinical_variables: variables in ADNI dataset
- pca_components: column related to PCA components

After fix input data path, we set number of partitions and partitions save directory.

```{r, eval=FALSE}
wgs_adno_partitioner(
  partition_number = 300,
  clinical_variables = c("MMSE", "GENDER","AGE", "MMSE.bl", "PTEDUCAT", "PTID", "VISID"),
  pca_components = c("PC1", "PC2", "PC3"),
  wgs_path = "/data/WGScompletedQC_phenoed.raw",
  pca_path = "/data/ADNIprunedpostQCPCA.eigenvec",
  adni_path = "/data/adnimerge.RData",
  partitions_save_path = "/data/ADNI/data_partitions/")
```

We need a R script that runs on each core to execute LMER modelling in parallel.
`function_to_Rscript` helps to generate an R script from a function in a package or
source from the local code. This script is designed to run independently on each core.
To create script we need set:

- `script_name`: the name of the created script.
- `function_name`: the name of a function in a package or the address of a function's source file.
- `packages`: list of packages that are loaded in the script.
- `arguments`: includes function input arguments
- `arguments_class`: contains a vector of argument types (character, integer, numeric).

```{r, eval=FALSE}
function_to_Rscript(
  script_name = "/scripts/Parallel_Modeling.R", 
  function_from_source = NULL,
  function_name = "lmer_modeling",
  packages =  c("HLSGUtils"),
  arguments = c("data_path", "simulation_name", "formula", "save_model_path"),
  arguments_class = c("character", "character", "character", "character")
)
```

`lmer_modeling` function is written to run LMER model on each data partitions.
It needs :

- `data_path`: the path of partition data, 
- `formula`: lmer formula that contains random effect term, 
- `simulation_name`: is used in the save model file name.
- `save_model_path`: The directory of saving model output.

Finally we use parallel_rscripts to run scripts on parallel cores.
The main arguments of function are:


- `rscript_path`:	path to the script that is run concurrently.
- `args`: script's input arguments.
- `free_memory_treshold`:	upper bound on memory usage percentage
- `free_cpu_treshold`: upper bound on CPU percentage

```{r,eval=FALSE}
library(HLSGUtils)

# `lmer_modeling` input arguments
partitions_files = list.files("/data/data_partitions/", full.names = T)
save_model_path = "/data/models/"
formula = paste0("'","MMSE~GENDER+AGE+MMSE.bl+PC1+PC2+PC3+PTEDUCAT+copy_number+(1|PTID)+(1|VISID)","'")
simulation_name = "full_model"

parallel_rscripts(
  rscript_path = "/scripts/Parallel_Modeling.R",
  args = list(data_path = partitions_files,
              simulation_name = simulation_name,
              formula = formula,
              save_model_path = save_model_path),
  free_memory_treshold = 80,
  free_cpu_treshold = 80,
  sleep_time = 10
)
```

To aggregate models coefficients use `aggregate_coefficients` function.

```{r, eval=FALSE}
aggregate_coefficients(
  save_model_directory = "/data/models/",
  model_names_pattern = "full_model",
  save_model_path = "/data/aggregated_models/full_model.rds"
)
```


